---
title: "Statistical Inference"
output: html_document
---

# Statistical Inference 


# This Notebook contains rough calculations only. Nothing is finalized or clean. It was a scratchpad.

To better inform our mathematical models, statistical techniques can be deployed for parameter estimation.

A collection of varied datasets were collected from WHO to give a good indication of the state of Malaria in the $3$ target countries:
  - Nigeria
  - DRC
  - Uganda
  
We wish to look at the relationship in incidence growth as a function of social-economic factors. The analysis can easily be extended to include other countries.


## Clean dataset

The various datasets need to be collected & aggregated into one dataframe. Most of the variables have a time component & as such we need to store the data in a _panel data_ fashion.

```{r, setup, include=FALSE}

setwd("~/Desktop/MSc Advanced Analytics/MMID/Assignment")

# All available datasets
dirs <- list.dirs('~/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets', full.names=T)




# ________________________________________ Social Economic Datasets ________________________________________ x


# -----  Malaria (directly related) -----x

percent_children_u5_w_fever_recieving_antimalarial_drugs <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/%children under 5 with fever receiving antimalarial drugs/data.csv", header = T, sep=';')

use_of_insecticide_treated_bednets_percentage_of_pop_under_5 <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/use of insecticide-treated bed nets (% of under 5 population)/API_SH.MLR.NETS.ZS_DS2_en_csv_v2_1139380.csv", skip=3)

incidence_of_malaria_per_1000_pop_at_risk <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/incidence of malaria per 1000 population at risk/API_SH.MLR.INCD.P3_DS2_en_csv_v2_1124271.csv", skip = 3)



# ----- Education -----x

literacy_rate_adults_above_15 <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/literacy rate adults 15 & above/API_SE.ADT.LITR.ZS_DS2_en_csv_v2_1120965.csv", skip = 3)

net_secondary_school_enrollment <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/net secondary school enrollment/API_SE.SEC.NENR_DS2_en_csv_v2_1131166.csv", skip=3)

researchers_in_RD_per_million_people <- read.csv( "/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/researchers in R&D (per million people)/API_SP.POP.SCIE.RD.P6_DS2_en_csv_v2_1132073.csv", skip = 3)

no_scientific_techinical_journal_articles <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/scientific & technical journal articles/API_IP.JRN.ARTC.SC_DS2_en_csv_v2_1124961.csv", skip=3)




# ----- Economic Growth & Dev -----x

electric_power_consumption_kwh_per_capital <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/electric power consumption (kwh per capita)/data.csv", sep=';')

GPD <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/GDP data/data.csv", sep=';')

GPD_per_capita <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/GDP per capita/data.csv", sep=';')




# ----- Inequality -----x

gini_coeff <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/Gini coeff/API_SI.POV.GINI_DS2_en_csv_v2_1120931.csv", sep=',', header = T, skip = 3)

income_share_held_by_lowest_10_percent <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/income share held by lowest 10%/API_SI.DST.FRST.10_DS2_en_csv_v2_1122524.csv", skip = 3)




# ----- Population -----x

net_migration <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/net_migration/API_SM.POP.NETM_DS2_en_csv_v2_1120954.csv", skip=3)

total_population <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/population total/API_SP.POP.TOTL_DS2_en_csv_v2_1120881.csv", skip = 3)

population_growth <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/population_growth/API_SP.POP.GROW_DS2_en_csv_v2_1122793.csv", skip = 3)

rural_population_percentage <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/rural_population_percentage/API_SP.RUR.TOTL.ZS_DS2_en_csv_v2_1126704.csv", skip=3)



# new - climate

CO2_emissions <- read.csv("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/climate change CO2 emissions (kt)/API_19_DS2_en_csv_v2_1125137.csv", skip=3)



```

We wish to visualize the trends in the social economic data, the function below allows one to to specify a dataset & a range of countries & a visualization is produced.


```{r}
library(tidyverse)

vis <- function(countries, dataset, title, ylab, standardize_as_percentage=F) {
  cols <- paste('X', '1960':'2019', sep = '')
  names(dataset)[names(dataset) %in% cols] <- '1960':'2019'
  dataset_long <- gather(dataset, year, count, `1960`:`2019`)[,c('Country.Name', 'year', 'count')]
  dataset_long$year <- as.Date(dataset_long$year, '%Y')
  
  if (standardize_as_percentage) {
    dataset_long <- dataset_long %>% group_by(Country.Name) %>% 
      mutate(count = (count-min(count, na.rm=T))/max(count, na.rm=T))
  }
  
  remove_missing(dataset_long[dataset_long$Country.Name %in% countries, ]) %>%
  ggplot(aes(x=year, y=count, col=Country.Name, na.rm=T)) + geom_line() + theme_minimal() + ylab(ylab) + 
  theme(plot.title = element_text(hjust = .5)) + ggtitle(title)
}


# hyper parameters
countries <- c('Nigeria', "Democratic Republic of the Congo", "Uganda", 'Congo, Dem. Rep.')



# ----- Population -----x

# population
vis(countries, dataset=total_population, title = 'Total Population', ylab = '', standardize_as_percentage = T)

# population growth
vis(countries, dataset=population_growth, title = 'Population Growth', ylab = '', standardize_as_percentage = F)

# net migration
vis(countries, dataset=net_migration, title = 'Net Migration', ylab = '', standardize_as_percentage = F)

# rural population
vis(countries, dataset=rural_population_percentage, title = '% Rural Population', ylab = '', standardize_as_percentage = F)
 
  


# ----- Inequality -----x

# gini coefficient
vis(countries, dataset=gini_coeff, title = 'Gini Coefficient', ylab = '', standardize_as_percentage = F)

# income inequality
vis(countries, dataset=income_share_held_by_lowest_10_percent, title = 'Income percentage held by lowest 10%', ylab = '', standardize_as_percentage = F)





# -----  Malaria (directly related) -----x

# percentage children u5 with fever receiving antimalarial drugs 
vis(countries, dataset=percent_children_u5_w_fever_recieving_antimalarial_drugs, title = '% children under 5 with fever receiving antimalarial drugs ', ylab = '', standardize_as_percentage = F)

# use of ITNs
vis(countries, dataset=use_of_insecticide_treated_bednets_percentage_of_pop_under_5, 
    title='% children under 5 that use insecticide treated bednets', ylab = '', standardize_as_percentage = F)
 
# incidence of malaria
vis(countries, dataset=incidence_of_malaria_per_1000_pop_at_risk, 
    title='Incidence of malaria per 1000 people at risk', ylab = '', standardize_as_percentage = F)
 



# ----- Education -----x

# adult literact 
vis(countries, dataset=literacy_rate_adults_above_15, 
    title='Literacy rate (adults above 15)', ylab = '', standardize_as_percentage = F)

# secondary school enrollment
vis(countries, dataset=net_secondary_school_enrollment, 
    title='Secondary school enrollment', ylab = '', standardize_as_percentage = F)
 
# no. researchers
vis(countries, dataset=researchers_in_RD_per_million_people, 
    title='Researchers in R&D (per million people)', ylab = '', standardize_as_percentage = F)

# scientific output
vis(countries, dataset=no_scientific_techinical_journal_articles, 
    title='No. of Scientific or Technical Journals Written', ylab = '', standardize_as_percentage = F)






# ----- Economic Growth & Dev -----x

# electicity use
vis(countries, dataset=electric_power_consumption_kwh_per_capital, 
    title='Electric Power Consumption (kwh/capita)', ylab = '', standardize_as_percentage = F)


# GDP
vis(countries, dataset=GPD, title='GDP', ylab = '', standardize_as_percentage = F)
vis(countries, dataset=GPD, title='Standardized GDP', ylab = '', standardize_as_percentage = T)


# GDP per capita
vis(countries, dataset=GPD_per_capita, title='GDP per Capita', ylab = '', standardize_as_percentage = F)
vis(countries, dataset=GPD_per_capita, title='Standardized GDP per Capita', ylab = '', standardize_as_percentage = T)







```


## Malaria incidence data

Here we process the cases number statistics.


```{r}

# ---- malaria stats ----x

estimated_cases <- read.csv('../../datasets/malaria statistics/Estimated_cases.csv', header = F)
names(estimated_cases) <- trimws(estimated_cases[2,])
estimated_cases <- estimated_cases[-c(1,2),]



estimated_deaths <- read.csv('../../datasets/malaria statistics/Estimated_deaths.csv', header = F)
names(estimated_deaths) <- trimws(estimated_deaths[2,])
estimated_deaths <- estimated_deaths[-c(1,2),]


prevention <- read.csv('../../datasets/malaria statistics/Prevention_children_under_ITNs.csv', header = F)
names(prevention) <- trimws(prevention[1,])
prevention <- prevention[-1,]

```


## Estimated Cases

Visualize the case data

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)

countries <- c('Nigeria', "Democratic Republic of the Congo", "Uganda")


numeric_column <- function(col) {
  col %>% 
  gsub(pattern = '\\[.*]', replacement = "") %>%
  gsub(pattern = ' ', replacement = '') %>%
  as.numeric()
}


# ---- estimated cases ----
estimated_cases[,-1] <- apply(estimated_cases[,-1], 2, FUN = numeric_column)

long_case_data <- gather(estimated_cases, year, cases, `2017`:`2010`, factor_key=TRUE)
long_case_data$year <- as.Date(as.character(long_case_data$year), '%Y')

long_case_data[long_case_data$Country %in% countries, ] %>%
  ggplot(aes(x=year, y=cases, col=Country)) + geom_line() + theme_minimal() + 
  theme(plot.title = element_text(hjust = .5)) + ggtitle('Malaria Cases')



# ---- estimated deaths ----
estimated_deaths[,-1] <- apply(estimated_deaths[,-1], 2, FUN=numeric_column)

long_death_data <- gather(estimated_deaths, year, deaths, `2017`:`2010`, factor_key=TRUE)
long_death_data$year <- as.Date(as.character(long_death_data$year), '%Y')

long_death_data[long_death_data$Country %in% countries, ] %>%
  ggplot(aes(x=year, y=deaths, col=Country)) + geom_line() + theme_minimal() + 
  theme(plot.title = element_text(hjust = .5)) + ggtitle('Malaria Deaths')



# ---- Prevention: children under ITNs ----
prevention[,-c(1,2)] <- apply(matrix(prevention[,-c(1,2)]), 2, FUN=numeric_column)

long_death_data <- gather(estimated_deaths, year, deaths, `2017`:`2010`, factor_key=TRUE)
long_death_data$year <- as.Date(as.character(long_death_data$year), '%Y')




# clean
compress_date_range <- function(index) {
  splt <- strsplit(prevention$Year[index], '-')
  round(mean(lapply(splt, as.numeric)[[1]]))
}

# indeces of date ranges 
ind <- grep('-', prevention$Year)
prevention$Year[ind] <- unlist(lapply(ind, FUN = compress_date_range))
prevention$Year <- as.Date(as.character(prevention$Year), '%Y')

prevention$child <- prevention$`Children aged &lt;5 years sleeping under insecticide-treated nets (%)`

prevention[prevention$Country %in% countries, ] %>%
  ggplot(aes(x=Year, y=child, col=Country)) + geom_line() + theme_minimal() + 
  theme(plot.title = element_text(hjust = .5)) + ggtitle('Percentage of children under 5 sleeping under ITNs')
```







----------------------------------------------------------------------------------------------------------------------------------------------


# Model

Clean Matrices

```{r}




library(dplyr)
library(ggplot2)
library(tidyverse)


# ---------- Clean Data ----------x
clean_data <- function(dataset, standardize_as_percentage=F) {
  cols <- paste('X', '1960':'2019', sep = '')
  names(dataset)[names(dataset) %in% cols] <- '1960':'2019'
  dataset_long <- gather(dataset, year, count, `1960`:`2019`)[,c('Country.Name', 'year', 'count')]
  
  if (standardize_as_percentage) {
    dataset_long <- dataset_long %>% group_by(Country.Name) %>% 
      mutate(count = (count-min(count, na.rm=T))/max(count, na.rm=T))
  }
  
  dataset_long
}


# ---------- Create Response Variable ----------x

Y <- clean_data(incidence_of_malaria_per_1000_pop_at_risk)
Y <- Y[!is.na(Y$count),]
countries_list <- Y$Country.Name







# ---------- Create Independent Variable ----------x
perc_ITNs_pop_under_5 <- clean_data(use_of_insecticide_treated_bednets_percentage_of_pop_under_5)
perc_under_5_rec_antimal_drugs <- clean_data(percent_children_u5_w_fever_recieving_antimalarial_drugs)
std_gdp <- clean_data(GPD, standardize_as_percentage = T)
rural_pop_perc <- clean_data(rural_population_percentage)
lit_rate_adults <- clean_data(literacy_rate_adults_above_15)


```

Next the data needs to be cleaned, missing values imputed as averages & all data concatonated into one dataset.

```{r}


citation("lme4")
library(imputeTS)


# ---- remove countries without a response ----x
prune_countries <- function(data, countries=countries_list) {
  data[data$Country.Name %in% countries,]
}

perc_ITNs_pop_under_5 <- prune_countries(perc_ITNs_pop_under_5)
perc_under_5_rec_antimal_drugs <- prune_countries(perc_under_5_rec_antimal_drugs)
std_gdp <- prune_countries(std_gdp)
rural_pop_perc <- prune_countries(rural_pop_perc)
lit_rate_adults <- prune_countries(lit_rate_adults)



# ----- impute NA's as moving average -----x
lit_rate_adults <- imputeTS::na_ma(lit_rate_adults)
perc_ITNs_pop_under_5 <- na_ma(perc_ITNs_pop_under_5)
perc_under_5_rec_antimal_drugs <- na_ma(perc_under_5_rec_antimal_drugs)
std_gdp <- na_ma(std_gdp)
rural_pop_perc <- na_ma(rural_pop_perc)




# ----- compute GDP growth rate -----x
std_gdp <- std_gdp %>% group_by(Country.Name) %>% mutate(grw_rate = (count-lag(count))/lag(count))
std_gdp$grw_rate[std_gdp$grw_rate==Inf] <- 0
std_gdp <- na_ma(std_gdp)



# ---------- X: Design Matrix ----------x


# --- fix col names ---x
names(lit_rate_adults) <- c("Country.Name", "year", "lit_rate_adults")
names(perc_ITNs_pop_under_5) <- c("Country.Name", "year", "perc_ITNs_pop_under_5")
names(perc_under_5_rec_antimal_drugs) <- c("Country.Name", "year", "perc_under_5_rec_antimal_drugs")
names(std_gdp) <- c("Country.Name", "year", "std_gdp", 'std_gdp_growth_rate')
names(rural_pop_perc) <- c("Country.Name", "year", "rural_pop_perc")


X <- Reduce(function(x, y) merge(x, y, by=c('Country.Name', 'year')), list(lit_rate_adults, perc_ITNs_pop_under_5, perc_under_5_rec_antimal_drugs, std_gdp, rural_pop_perc))
y <- Y 

dim(y)  
dim(X)

unique(y$year)
unique(X$year)

# only use data over Y's range (years)
X <- X[X$year %in% y$year,]

dim(y)  
dim(X)


# ---- convert to date type ----x 
y$year <- as.Date(y$year, '%Y')
X$year <- as.Date(X$year, '%Y')

# ----- concate data ------x
names(y) <- c('Country.Name', 'year', 'mal_inc_per_1000_at_risk')

data <- merge(X, y, by=c('Country.Name', 'year'))


# ---- save dataframe! ----x
save(data, file="/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/clean_model_data.Rda")
save(countries_list, file="/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/countries_list.Rda")


load("/Users/zachwolpe/Desktop/MSc Advanced Analytics/MMID/Assignment/datasets/clean_model_data.Rda")

countries_list

```


# Model: Heirarchical GLM via REML

Now that the data is clean we need to formulate a model. There is a nature temperal structure in the data, fitting a standard GLM is thus inappropriate as it does not dependent covariate structure.

To capture the heirarchical GLM in an unbiased way, where the temporal element is considered by considering an individual a heirarchy (group) by which elements are


```{r}

# ---------- Fit Mixed Effect Model: Hierarchy==Time ----------x

library(lme4)



GLMM <- lmer(mal_inc_per_1000_at_risk ~ lit_rate_adults + perc_ITNs_pop_under_5 + perc_under_5_rec_antimal_drugs  + 
               std_gdp_growth_rate + rural_pop_perc + (year|Country.Name), data = data)




summary(GLMM)

yhat <- predict(GLMM)
arg_data <- cbind(data, yhat)
countries_list <- unique(data$Country.Name)


plot_for_countries <- function() {
    cc <- sample(countries_list, 1)
    arg_data[arg_data$Country.Name==cc,] %>% 
    ggplot(aes(x=year, y=mal_inc_per_1000_at_risk)) + geom_point(col='darkred') + geom_line(col='darkred') +
    geom_point(aes(x=year, y=yhat), col='steelblue') + geom_line(aes(x=year, y=yhat), col='steelblue') + ggtitle(cc) +
      ylab('Malaria Incidence (per 1000 people at Risk)') + xlab('Year') +  theme_minimal() + 
      theme(plot.title =element_text(hjust = 0.5))
    

}

plot_for_countries()
  

```


# Sigmoid Mapping 

Fit the Sigmoid for a particular country & timeframe

Year Range: 2000-2018

Given model inputs! predict !  

```{r}

n_years <- 10
country <- 'World'







compute_new_prediction <- function(country) {
  
  
  
  # --- compute growth in each variable ---x
  sub <- arg_data[arg_data$Country.Name == country,]
  sub[sub==0] <- NA
  sub <- na_ma(sub)
  
  
  
  # ---- comptute growth rates ---- x
  sub <- sub %>% mutate(lit_rate_adults_gr = 
                          ifelse(is.na((lit_rate_adults-lag(lit_rate_adults))/lit_rate_adults), 1,
                                 (lit_rate_adults-lag(lit_rate_adults))/lit_rate_adults),
                        
                        perc_ITNs_pop_under_5_gr =
                          ifelse(is.na((perc_ITNs_pop_under_5-lag(perc_ITNs_pop_under_5))/lag(perc_ITNs_pop_under_5)), 1,
                                 (perc_ITNs_pop_under_5-lag(perc_ITNs_pop_under_5))/lag(perc_ITNs_pop_under_5)),
                        
                        perc_under_5_rec_antimal_drugs_gr = 
                          ifelse(is.na((perc_under_5_rec_antimal_drugs-lag(perc_under_5_rec_antimal_drugs))/lag(perc_under_5_rec_antimal_drugs)), 1,
                                 (perc_under_5_rec_antimal_drugs-lag(perc_under_5_rec_antimal_drugs))/lag(perc_under_5_rec_antimal_drugs)),
                        
                        std_gdp_gr = 
                          ifelse(is.na((std_gdp-lag(std_gdp))/lag(std_gdp)), 1,
                                 (std_gdp-lag(std_gdp))/lag(std_gdp)),
  
                        std_gdp_growth_rate_gr = 
                          ifelse(is.na((std_gdp_growth_rate-lag(std_gdp_growth_rate))/lag(std_gdp_growth_rate)), 1,
                                 (std_gdp_growth_rate-lag(std_gdp_growth_rate))/lag(std_gdp_growth_rate)),
  
                        rural_pop_perc_gr = 
                          ifelse(is.na((rural_pop_perc-lag(rural_pop_perc))/lag(rural_pop_perc)), 1,
                                 (rural_pop_perc-lag(rural_pop_perc))/lag(rural_pop_perc)))
  
  
  
  
  # ---- inititional variables ----x
  lit_rate_adults_2                 <- c(sub$lit_rate_adults[1] * 1)
  perc_ITNs_pop_under_5_2           <- c(sub$perc_ITNs_pop_under_5[1] * 1)
  perc_under_5_rec_antimal_drugs_2  <- c(sub$perc_under_5_rec_antimal_drugs[1] * 1)
  std_gdp_2                         <- c(sub$std_gdp[1] * 1)
  std_gdp_growth_rate_2             <- c(sub$std_gdp_growth_rate[1] * 1)
  rural_pop_perc_2                  <- c(sub$rural_pop_perc[1] * 1)
  
  
  # ---- compute new variables ----x
  for (i in 2:19) {
    lit_rate_adults_2 <- c(lit_rate_adults_2, lit_rate_adults_2[i-1] + lit_rate_adults_2[i-1]*sub$lit_rate_adults_gr[i])
    perc_ITNs_pop_under_5_2 <- c(perc_ITNs_pop_under_5_2, perc_ITNs_pop_under_5_2[i-1] + perc_ITNs_pop_under_5_2[i-1]*sub$perc_ITNs_pop_under_5_gr[i])
    perc_under_5_rec_antimal_drugs_2 <- c(perc_under_5_rec_antimal_drugs_2, perc_under_5_rec_antimal_drugs_2[i-1] +
                                            perc_under_5_rec_antimal_drugs_2[i-1]*sub$perc_under_5_rec_antimal_drugs_gr[i])
    std_gdp_2 <- c(std_gdp_2, std_gdp_2[i-1] + std_gdp_2[i-1]*sub$std_gdp_gr[i])
    std_gdp_growth_rate_2 <- c(std_gdp_growth_rate_2, std_gdp_growth_rate_2[i-1] + std_gdp_growth_rate_2[i-1]*sub$std_gdp_growth_rate_gr[i])
    rural_pop_perc_2 <- c(rural_pop_perc_2, rural_pop_perc_2[i-1] + rural_pop_perc_2[i-1]*sub$rural_pop_perc_gr[i])
  }
  
  
  # ---- create new dataset ----x 
  sub <- data.frame(Country.Name = sub$Country.Name, year = sub$year, lit_rate_adults=lit_rate_adults_2,
                    perc_ITNs_pop_under_5=perc_ITNs_pop_under_5_2, perc_under_5_rec_antimal_drugs=perc_under_5_rec_antimal_drugs_2,
                    std_gdp=std_gdp_2, std_gdp_growth_rate=std_gdp_growth_rate_2, rural_pop_perc=rural_pop_perc_2)
  
  
  
  sub$yhat <- predict(GLMM, sub)
  sub

}




sub <- compute_new_prediction('Uganda')

arg_data[arg_data$Country.Name=='Uganda',] %>% 
ggplot(aes(x=year, y=mal_inc_per_1000_at_risk)) + geom_point(col='darkred') + geom_line(col='darkred') +
geom_point(aes(x=year, y=yhat), col='pink') + 
  # geom_line(aes(x=year, y=yhat), col='steelblue') + ggtitle(cc) +
  ylab('Malaria Incidence (per 1000 people at Risk)') + xlab('Year') +  theme_minimal() + 
  geom_line(aes(x = sub$year, y=sub$yhat), col='lightblue') +
  theme(plot.title =element_text(hjust = 0.5))
    




```


# Sigmoid Transfermation 


```{r}


arg_data[arg_data$Country.Name=='Uganda',] %>% 
ggplot(aes(x=year, y=mal_inc_per_1000_at_risk)) + geom_point(col='darkred') + geom_line(col='darkred') +
geom_point(aes(x=year, y=yhat), col='pink') + 
  # geom_line(aes(x=year, y=yhat), col='steelblue') + ggtitle(cc) +
  ylab('Malaria Incidence (per 1000 people at Risk)') + xlab('Year') +  theme_minimal() + 
  geom_line(aes(x = sub$year, y=sub$yhat), col='lightblue') +
  theme(plot.title =element_text(hjust = 0.5))
    

arg_data[arg_data$Country.Name=='Uganda',][19,]


first_inc <- arg_data[arg_data$Country.Name=='Uganda',]$mal_inc_per_1000_at_risk[1]
last_inc <- arg_data[arg_data$Country.Name=='Uganda',]$mal_inc_per_1000_at_risk[19]

print(paste(round((last_inc-first_inc)/first_inc*100, 4), '% change', sep=''))


# given the models predicted change 
old_ch <- -43
new_ch <- -47

# the old should map to the prior, the prior is given.
prior_theta <- 3

# determine the range: theta: 1-4

# center around old == mean == zero

sigmoid <- function(x) 1/(1+exp(-x))

x <- seq(-10, 10, length.out = 100)

plot(x, sigmoid(x), type='l', main='standard sigmoid', frame=F, col='orange')


# x range need not be standardized

# given the models predicted change 
old_ch <- -.43
new_ch <- -.47

# y range needs to fit the parameter range: 1-4
plot(x, 1+(4-1)*sigmoid(x), type='l', main='standard sigmoid', frame=F, col='orange')
abline(v = old_ch-new_ch, h = 1+(4-1)*sigmoid(old_ch-new_ch))
```


# Temerature

```{r}
t<-seq(0,100,1)
beta1<-1
beta0<-1

#https://www.r-bloggers.com/bayesian-models-in-r-2/


likelihood <- function(t) {
  cos(t/2)*3  + cos(t)*3 + 28
}

plot(t, y, type="l", frame=F, main='Simulated Nigeria Temperature Data', col='steelblue', xlab='', ylab='temperature')


y <- likelihood(1:1000)



dy <- density(y)



x_range <- seq(15,45,length.out = 512) 
prior <- dnorm(sample(x_range, size = 512/2), mean=30, 0.2) + dnorm(sample(x_range, size = 512/2), mean=22, 0.2)





prior <- c(rnorm(512/2, mean=30, 1), rnorm(512/2, mean=33, 1))
dp <- density(prior)
posterior <- dy$y* dp$y


# --- likelihood ---x
plot(density(y), col='darkblue', ylab='density', xlab='Temperature', main='Posterior Distribution of Temperature', frame=F)
polygon(density(y), col=adjustcolor("darkblue",alpha.f=0.1), border=NA) 

# --- prior ---x
lines(dp, col='darkgreen', type='l')
polygon(dp, col=adjustcolor("darkgreen",alpha.f=0.1), border=NA) 

# --- posterior ---x
lines(x_range, posterior, col='steelblue', type='l')
abline(v=x_range[which(posterior==max(posterior))], col='black')
polygon(x_range, posterior, col=adjustcolor("steelblue",alpha.f=0.2), border=NA)

legend('topright', legend=c('density', 'prior', 'posterior'), col=c('darkblue', 'darkgreen', 'steelblue'), lty=1, cex=0.8)


```



# Bayesian Temperature

```{r}
rangeP <- seq(0, 1, length.out = 100)
plot(rangeP, dbinom(x = 8, prob = rangeP, size = 10),
type = "l", xlab = "P(Black)", ylab = "Density")

	
lines(rangeP, dnorm(x = rangeP, mean = .5, sd = .1) / 15,
col = "red")

lik <- dbinom(x = 8, prob = rangeP, size = 10)
prior <- dnorm(x = rangeP, mean = .5, sd = .1)
lines(rangeP, lik * prior, col = "green")




```




## Parasital Parameters

```{r}



plot_densities <- function(density, prior, posterior) {
  plot(dens, col='darkblue', ylab='density', xlab='Temperature', main='Posterior Distribution of phi', frame=F, type='l')
  polygon(dens, col=adjustcolor("darkblue",alpha.f=0.1), border=NA) 
  
  
  lines(prior, col='darkgreen', type='l')
  polygon(prior, col=adjustcolor("darkgreen",alpha.f=0.1), border=NA) 
  
  lines(posterior, col='darkred', type='l')
  abline(v=x_r[which(posterior==max(posterior))], col='black')
  polygon(posterior, col=adjustcolor("darkred",alpha.f=0.4), border=NA)
  legend('topright', legend=c('density', 'prior', 'posterior'), col=c('darkblue', 'darkgreen', 'darkred'), lty=1, cex=0.8)
}


# range 
x_r <- seq(20,42,length.out = 100)

# prior 
prior <- dunif(x_r,min=22, max=40)
# prior <- dnorm(x_r, mean = 22, sd = 2)


# density
dens <- dnorm(x_r, mean = 30, sd = 2)

# posterior  
posterior <- prior * dens
posterior <- posterior/sum(posterior)



plot_densities(dens, prior, posterior)


```





























